### 2022-06-21 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@eea6e201f3438c201940b84fa06464c6f96cb8a2 🚀

For every species pool, we measured its Local Contribution to
Beta-Diversity;\textsuperscript{58} LCBD works from a species-data
matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
and sites are columns, and a value of 1 indicates occurrence. We
extracted the Y matrix assuming that every pixel represents a unique
location, and following best practices\textsuperscript{59} transformed
it using Hellinger's distance to account for unequal bat richness at
different pixels. The correction of raw community data is particularly
important for two reasons: first, it prevents the artifact of richer
sites having higher importance; second, it removes the effect of overall
species richness, which is already incorporated in the phylogenetic
diversity component. High values of LCBD indicate that the pixel has a
community that is on average more dissimilar in species composition than
what is expected knowing the entire matrix, i.e.~a more unique
community. Recent results by\textsuperscript{60} shows that LCBD
measures are robust with regards to spatial scale, and are therefore
applicable at the global scale.

@@

To visualize the aggregated risk at the global scale, we combine the
three individual risk components (phylogenetic diversity, compositional
uniqueness, and viral sharing) using an additive color
model.\textsuperscript{62} In this approach, every risk component gets
assigned a component in the RGB color model (phylogenetic diversity is
green, compositional uniqueness is red, and viral sharing is blue). In
order to achieve a valid RGB measure, all components are re-scaled to
the {[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
diversity, and no compositional uniqueness is black, and a pixel with
maximal values for each is white. This additive model conveys both the
intensity of the overall risk, but also the nature of the risk as colors
diverge towards combinations of values for three risk components. Out of
the possible combinations, the most risky in terms or rapid
diversification and spillover potential is high phylogenetic diversity
and low viral sharing,\textsuperscript{63,64} in that this allows
multiple independent host-virus coevolutionary dynamics to take place in
the same location. In the colorimetric space, this correspond to yellow
-- because the HSV space is more amenable to calculations for feature
extraction,\textsuperscript{see \emph{e.g.} 65} we measured the risk
level by calculating the angular distance of the hue of each pixel to a
reference value of 60, and weighted this risk level by the value
component. Specifically, given a pixel with colorimetric coordinates
\((h,s,v)\), its ranged weighted risk value is

@@

To next represent phylogeography of betacoronaviruses in bats, we
aggregated and analyzed betacoronavirus sequence data. We used the
following query to pull all \emph{Betacoronavirus} sequence data from
the GenBank Nucleotide database except SARS-CoV-2;
(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
OR sars-cov-2{[}All Fields{]}). We added a single representative
sequence for SARS-CoV-2 and manually curated to remove sequences without
the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
indicating recombinant or laboratory strains including ``patent'',
``mutant'', ``GFP'', and ``recombinant''. We filtered over-represented
taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
betacoronavirus RdRp sequences were then aligned using MAFFT v 1.4.0
{[}\textsuperscript{66}; Algorithm FFT-NS-2, Scoring matrix 200PAM /
k=2, gap open penalty 1.53m offset value 0.123{]} and a maximum
likelihood tree reconstructed in IQ-TREE v 1.6.12\textsuperscript{67}
with ModelFinder\textsuperscript{68} ultrafast bootstrap
approximation\textsuperscript{69} with a general time reversible model
with empirical base frequencies and the 5-discrete-rate-category
FreeRaye model of nucleotide substitution (GTR+F+R5).

We first tested the hypothesis that hotspots of viral diversification
would track hotspots of bat diversification. To do so, we plotted the
number of known bat hosts (specifically only those included in the
phylogeny, so there was a 1:1 correspondence between data sources)
against the ``mean evolutionary distinctiveness'' of the associated
viruses. To calculate this, we derived the fair proportions evolutionary
distinctiveness\textsuperscript{70} for each of the viruses in the tree,
then averaged these at the bat species level, projected these values
onto their geographic distributions, and averaged across every bat found
in a given pixel. As such, this can be thought of as a map of the mean
evolutionary distinctiveness of the known viral community believed to be
associated with a particular subset of bats present.

@@

Subsequently, we tested the hypothesis that the biogeography of bat
betacoronaviruses should track the biogeography of their hosts. To test
this idea, we loosely adapted a method from,\textsuperscript{71,72} who
proposed a phylogenetic method for the delineation of animal
biogeographic regions. In their original method, a distance matrix -
where each row or column represents a geographic raster's grid cell, and
the dissimilarity values are the ``beta diversity similarity'' of their
community assemble - undergoes non-metric multidimensional scaling
(NMDS); the first two axes of the NMDS are projected geographically
using a four-color bivariate map. Here, we build on this idea with an
entirely novel methodology. First, we measure the phylogenetic distance
between the different viruses in the betacoronaviruses tree by using the
cophenetic function in \texttt{ape};\textsuperscript{73} subsequently,
we take a principal components analysis of that distance matrix (readily
interchangeable for NMDS in this case) to project the viral tree into an
n-dimensional space. We then take the first two principal components
and, as with the evolutionary distinctiveness analysis, aggregated these
to a mean host value and projected them using a four-color bivariate
map.

@@

\leavevmode\hypertarget{ref-Thompson2005GeoMos}{}%
\CSLLeftMargin{6. }
\CSLRightInline{Thompson, J. N. \emph{The Geographic Mosaic of
Coevolution}. (University Of Chicago Press, 2005).}

\leavevmode\hypertarget{ref-Thompson1994CoePro}{}%
\CSLLeftMargin{7. }
\CSLRightInline{Thompson, J. N. \emph{The Coevolutionary Process}.
(University of Chicago Press, 1994).}

\leavevmode\hypertarget{ref-Janzen1980WheIt}{}%
\CSLLeftMargin{8. }
\CSLRightInline{Janzen, D. H. When is it Coevolution? \emph{Evolution}
\textbf{34}, 611--612 (1980).}

\leavevmode\hypertarget{ref-Price2002MacThe}{}%
\CSLLeftMargin{9. }
\CSLRightInline{Price, P. W. \emph{Macroevolutionary Theory on
Macroecological Patterns}. (Cambridge University Press, 2002).
doi:\href{https://doi.org/10.1017/CBO9780511615030}{10.1017/CBO9780511615030}.}

\leavevmode\hypertarget{ref-Gomulkiewicz2007DosDon}{}%
\CSLLeftMargin{10. }
\CSLRightInline{Gomulkiewicz, R. \emph{et al.} Dos and don'ts of testing
the geographic mosaic theory of coevolution. \emph{Heredity}
\textbf{98}, 249--258 (2007).}

\leavevmode\hypertarget{ref-Thrall2007CoeSym}{}%
\CSLLeftMargin{11. }
\CSLRightInline{Thrall, P. H., Hochberg, M. E., Burdon, J. J. \& Bever,
J. D. Coevolution of symbiotic mutualists and parasites in a community
context. \emph{Trends in Ecology \& Evolution} \textbf{22}, 120--126
(2007).}

\leavevmode\hypertarget{ref-Drexler2014EcoEvo}{}%
\CSLLeftMargin{12. }
\CSLRightInline{Drexler, J. F., Corman, V. M. \& Drosten, C. Ecology,
evolution and classification of bat coronaviruses in the aftermath of
SARS. \emph{Antiviral Research} \textbf{101}, 45--56 (2014).}

\leavevmode\hypertarget{ref-Calisher2006BatImp}{}%
\CSLLeftMargin{13. }
\CSLRightInline{Calisher, C. H., Childs, J. E., Field, H. E., Holmes, K.
V. \& Schountz, T. Bats: Important Reservoir Hosts of Emerging Viruses.
\emph{Clinical Microbiology Reviews} \textbf{19}, 531--545 (2006).}

\leavevmode\hypertarget{ref-Moratelli2015BatZoo}{}%
\CSLLeftMargin{14. }
\CSLRightInline{Moratelli, R. \& Calisher, C. H. Bats and zoonotic
viruses: Can we confidently link bats with emerging deadly viruses?
\emph{Memórias do Instituto Oswaldo Cruz} \textbf{110}, 1--22 (2015).}

\leavevmode\hypertarget{ref-Mollentze2020VirZoo}{}%
\CSLLeftMargin{15. }
\CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
homogenous among taxonomic orders of mammalian and avian reservoir
hosts. \emph{Proceedings of the National Academy of Sciences}
\textbf{117}, 9423--9430 (2020).}

\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
\CSLLeftMargin{16. }
\CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
cross-species viral transmission risk. \emph{Nature} 1--1 (2022)
doi:\href{https://doi.org/10.1038/s41586-022-04788-w}{10.1038/s41586-022-04788-w}.}

\leavevmode\hypertarget{ref-Springer2013PhyBat}{}%
\CSLLeftMargin{17. }
\CSLRightInline{Springer, M. S. Phylogenetics: Bats United, Microbats
Divided. \emph{Current Biology} \textbf{23}, R999--R1001 (2013).}

\leavevmode\hypertarget{ref-Teeling2005MolPhy}{}%
\CSLLeftMargin{18. }
\CSLRightInline{Teeling, E. C. \emph{et al.} A Molecular Phylogeny for
Bats Illuminates Biogeography and the Fossil Record. \emph{Science (New
York, N.Y.)} \textbf{307}, 580--584 (2005).}

\leavevmode\hypertarget{ref-Villalobos2010DivFie}{}%
\CSLLeftMargin{19. }
\CSLRightInline{Villalobos, F. \& Arita, H. T. The diversity field of
New World leaf-nosed bats (Phyllostomidae). \emph{Global Ecology and
Biogeography} \textbf{19}, 200--211 (2010).}

\leavevmode\hypertarget{ref-Becker2022OptPre}{}%
\CSLLeftMargin{20. }
\CSLRightInline{Becker, D. J. \emph{et al.} Optimising predictive models
to prioritise viral discovery in zoonotic reservoirs. \emph{The Lancet
Microbe} (2022)
doi:\href{https://doi.org/10.1016/S2666-5247(21)00245-7}{10.1016/S2666-5247(21)00245-7}.}

\leavevmode\hypertarget{ref-Allen2017GloHot}{}%
\CSLLeftMargin{21. }
\CSLRightInline{Allen, T. \emph{et al.} Global hotspots and correlates
of emerging zoonotic diseases. \emph{Nature Communications} \textbf{8},
(2017).}

\leavevmode\hypertarget{ref-Olival2017HosVir}{}%
\CSLLeftMargin{22. }
\CSLRightInline{Olival, K. J. \emph{et al.} Host and viral traits
predict zoonotic spillover from mammals. \emph{Nature} \textbf{546},
646--650 (2017).}

\leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
\CSLLeftMargin{23. }
\CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
virome diversity in bats. \emph{Current Opinion in Virology}
\textbf{52}, 192--202 (2022).}

\leavevmode\hypertarget{ref-Letko2020BatVir}{}%
\CSLLeftMargin{24. }
\CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
K. \& Munster, V. J. Bat-borne virus diversity, spillover and emergence.
\emph{Nature Reviews Microbiology} \textbf{18}, 461--471 (2020).}

\leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
\CSLLeftMargin{25. }
\CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
Systems of Bats. \emph{Frontiers in Immunology} \textbf{11}, 26 (2020).}

\leavevmode\hypertarget{ref-Gorbunova2020WorGoe}{}%
\CSLLeftMargin{26. }
\CSLRightInline{Gorbunova, V., Seluanov, A. \& Kennedy, B. K. The World
Goes Bats: Living Longer and Tolerating Viruses. \emph{Cell Metabolism}
\textbf{32}, 31--43 (2020).}

\leavevmode\hypertarget{ref-Irving2021LesHos}{}%
\CSLLeftMargin{27. }
\CSLRightInline{Irving, A. T., Ahn, M., Goh, G., Anderson, D. E. \&
Wang, L.-F. Lessons from the host defences of bats, a unique viral
reservoir. \emph{Nature} \textbf{589}, 363--370 (2021).}

\leavevmode\hypertarget{ref-Jebb2020SixRef}{}%
\CSLLeftMargin{28. }
\CSLRightInline{Jebb, D. \emph{et al.} Six reference-quality genomes
reveal evolution of bat adaptations. \emph{Nature} \textbf{583},
578--584 (2020).}

\leavevmode\hypertarget{ref-MorenoSantillan2021LarGen}{}%
\CSLLeftMargin{29. }
\CSLRightInline{Moreno Santillán, D. D. \emph{et al.} Large-scale genome
sampling reveals unique immunity and metabolic adaptations in bats.
\emph{Molecular Ecology} mec.16027 (2021)
doi:\href{https://doi.org/10.1111/mec.16027}{10.1111/mec.16027}.}

\leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
\CSLLeftMargin{30. }
\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
distribution of bat hosts of sarbecoviruses: Implications for
conservation and public health. \emph{Proceedings of the Royal Society
B: Biological Sciences} \textbf{289}, 20220397 (2022).}

\leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
\CSLLeftMargin{31. }
\CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
SARS-Related Coronavirus Infection in Humans, China. \emph{Virologica
Sinica} \textbf{33}, 104--107 (2018).}

\leavevmode\hypertarget{ref-Ramshaw2019DatGeo}{}%
\CSLLeftMargin{32. }
\CSLRightInline{Ramshaw, R. E. \emph{et al.} A database of geopositioned
Middle East Respiratory Syndrome Coronavirus occurrences.
\emph{Scientific Data} \textbf{6}, 318 (2019).}

\leavevmode\hypertarget{ref-Lednicky2021IsoNov}{}%
\CSLLeftMargin{33. }
\CSLRightInline{Lednicky, J. A. \emph{et al.} Isolation of a Novel
Recombinant Canine Coronavirus from a Visitor to Haiti: Further Evidence
of Transmission of Coronaviruses of Zoonotic Origin to Humans.
\emph{Clinical Infectious Diseases: An Official Publication of the
Infectious Diseases Society of America} ciab924 (2021)
doi:\href{https://doi.org/10.1093/cid/ciab924}{10.1093/cid/ciab924}.}

\leavevmode\hypertarget{ref-Vlasova2022AniAlp}{}%
\CSLLeftMargin{34. }
\CSLRightInline{Vlasova, A. N. \emph{et al.} Animal alphacoronaviruses
found in human patients with acute respiratory illness in different
countries. \emph{Emerging Microbes \& Infections} \textbf{11}, 699--702
(2022).}

\leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
\CSLLeftMargin{35. }
\CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
Sister Species of Eidolon (Pteropodidae) with Evidence for Widespread
Panmixia. \emph{Acta Chiropterologica} \textbf{16}, 279--292 (2014).}

\leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
\CSLLeftMargin{36. }
\CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
Sequences From Malagasy Fruit Bats Define a Unique Evolutionary History
for This Coronavirus Clade. \emph{Frontiers in Public Health}
\textbf{10}, (2022).}

\leavevmode\hypertarget{ref-Hazarie2021IntPop}{}%
\CSLLeftMargin{37. }
\CSLRightInline{Hazarie, S., Soriano-Paños, D., Arenas, A.,
Gómez-Gardeñes, J. \& Ghoshal, G. Interplay between population density
and mobility in determining the spread of epidemics in cities.
\emph{Communications Physics} \textbf{4}, 191 (2021).}

\leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
\CSLLeftMargin{38. }
\CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
Origin in China. \emph{Emerging Infectious Diseases} \textbf{10},
1030--1037 (2004).}

\leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
\CSLLeftMargin{39. }
\CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
Becker, D. J. Sampling strategies and pre-pandemic surveillance gaps for
bat coronaviruses. 2022.06.15.496296 (2022)
doi:\href{https://doi.org/10.1101/2022.06.15.496296}{10.1101/2022.06.15.496296}.}

\leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
\CSLLeftMargin{40. }
\CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
synthesis of ecological and evolutionary determinants of bat diversity
across spatial scales. \emph{BMC ecology} \textbf{18}, 18 (2018).}

\leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
\CSLLeftMargin{41. }
\CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
World: A taxonomic and geographic database. \url{https://batnames.org/}
(2020).}

>>>>

For every species pool, we measured its Local Contribution to
Beta-Diversity;\textsuperscript{59} LCBD works from a species-data
matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
and sites are columns, and a value of 1 indicates occurrence. We
extracted the Y matrix assuming that every pixel represents a unique
location, and following best practices\textsuperscript{60} transformed
it using Hellinger's distance to account for unequal bat richness at
different pixels. The correction of raw community data is particularly
important for two reasons: first, it prevents the artifact of richer
sites having higher importance; second, it removes the effect of overall
species richness, which is already incorporated in the phylogenetic
diversity component. High values of LCBD indicate that the pixel has a
community that is on average more dissimilar in species composition than
what is expected knowing the entire matrix, i.e.~a more unique
community. Recent results by\textsuperscript{61} shows that LCBD
measures are robust with regards to spatial scale, and are therefore
applicable at the global scale.

@@

To visualize the aggregated risk at the global scale, we combine the
three individual risk components (phylogenetic diversity, compositional
uniqueness, and viral sharing) using an additive color
model.\textsuperscript{63} In this approach, every risk component gets
assigned a component in the RGB color model (phylogenetic diversity is
green, compositional uniqueness is red, and viral sharing is blue). In
order to achieve a valid RGB measure, all components are re-scaled to
the {[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
diversity, and no compositional uniqueness is black, and a pixel with
maximal values for each is white. This additive model conveys both the
intensity of the overall risk, but also the nature of the risk as colors
diverge towards combinations of values for three risk components. Out of
the possible combinations, the most risky in terms or rapid
diversification and spillover potential is high phylogenetic diversity
and low viral sharing,\textsuperscript{64,65} in that this allows
multiple independent host-virus coevolutionary dynamics to take place in
the same location. In the colorimetric space, this correspond to yellow
-- because the HSV space is more amenable to calculations for feature
extraction,\textsuperscript{see \emph{e.g.} 66} we measured the risk
level by calculating the angular distance of the hue of each pixel to a
reference value of 60, and weighted this risk level by the value
component. Specifically, given a pixel with colorimetric coordinates
\((h,s,v)\), its ranged weighted risk value is

@@

To next represent phylogeography of betacoronaviruses in bats, we
aggregated and analyzed betacoronavirus sequence data. We used the
following query to pull all \emph{Betacoronavirus} sequence data from
the GenBank Nucleotide database except SARS-CoV-2;
(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
OR sars-cov-2{[}All Fields{]}). We added a single representative
sequence for SARS-CoV-2 and manually curated to remove sequences without
the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
indicating recombinant or laboratory strains including ``patent'',
``mutant'', ``GFP'', and ``recombinant''. We filtered over-represented
taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
betacoronavirus RdRp sequences were then aligned using MAFFT v 1.4.0
{[}\textsuperscript{67}; Algorithm FFT-NS-2, Scoring matrix 200PAM /
k=2, gap open penalty 1.53m offset value 0.123{]} and a maximum
likelihood tree reconstructed in IQ-TREE v 1.6.12\textsuperscript{68}
with ModelFinder\textsuperscript{69} ultrafast bootstrap
approximation\textsuperscript{70} with a general time reversible model
with empirical base frequencies and the 5-discrete-rate-category
FreeRaye model of nucleotide substitution (GTR+F+R5).

We first tested the hypothesis that hotspots of viral diversification
would track hotspots of bat diversification. To do so, we plotted the
number of known bat hosts (specifically only those included in the
phylogeny, so there was a 1:1 correspondence between data sources)
against the ``mean evolutionary distinctiveness'' of the associated
viruses. To calculate this, we derived the fair proportions evolutionary
distinctiveness\textsuperscript{71} for each of the viruses in the tree,
then averaged these at the bat species level, projected these values
onto their geographic distributions, and averaged across every bat found
in a given pixel. As such, this can be thought of as a map of the mean
evolutionary distinctiveness of the known viral community believed to be
associated with a particular subset of bats present.

@@

Subsequently, we tested the hypothesis that the biogeography of bat
betacoronaviruses should track the biogeography of their hosts. To test
this idea, we loosely adapted a method from,\textsuperscript{72,73} who
proposed a phylogenetic method for the delineation of animal
biogeographic regions. In their original method, a distance matrix -
where each row or column represents a geographic raster's grid cell, and
the dissimilarity values are the ``beta diversity similarity'' of their
community assemble - undergoes non-metric multidimensional scaling
(NMDS); the first two axes of the NMDS are projected geographically
using a four-color bivariate map. Here, we build on this idea with an
entirely novel methodology. First, we measure the phylogenetic distance
between the different viruses in the betacoronaviruses tree by using the
cophenetic function in \texttt{ape};\textsuperscript{74} subsequently,
we take a principal components analysis of that distance matrix (readily
interchangeable for NMDS in this case) to project the viral tree into an
n-dimensional space. We then take the first two principal components
and, as with the evolutionary distinctiveness analysis, aggregated these
to a mean host value and projected them using a four-color bivariate
map.

@@

\leavevmode\hypertarget{ref-Geoghegan2017ComAna}{}%
\CSLLeftMargin{6. }
\CSLRightInline{Geoghegan, J. L., Duchêne, S. \& Holmes, E. C.
Comparative analysis estimates the relative frequencies of co-divergence
and cross-species transmission within viral families. \emph{PLOS
Pathogens} \textbf{13}, e1006215 (2017).}

\leavevmode\hypertarget{ref-Thompson2005GeoMos}{}%
\CSLLeftMargin{7. }
\CSLRightInline{Thompson, J. N. \emph{The Geographic Mosaic of
Coevolution}. (University Of Chicago Press, 2005).}

\leavevmode\hypertarget{ref-Thompson1994CoePro}{}%
\CSLLeftMargin{8. }
\CSLRightInline{Thompson, J. N. \emph{The Coevolutionary Process}.
(University of Chicago Press, 1994).}

\leavevmode\hypertarget{ref-Janzen1980WheIt}{}%
\CSLLeftMargin{9. }
\CSLRightInline{Janzen, D. H. When is it Coevolution? \emph{Evolution}
\textbf{34}, 611--612 (1980).}

\leavevmode\hypertarget{ref-Price2002MacThe}{}%
\CSLLeftMargin{10. }
\CSLRightInline{Price, P. W. \emph{Macroevolutionary Theory on
Macroecological Patterns}. (Cambridge University Press, 2002).
doi:\href{https://doi.org/10.1017/CBO9780511615030}{10.1017/CBO9780511615030}.}

\leavevmode\hypertarget{ref-Gomulkiewicz2007DosDon}{}%
\CSLLeftMargin{11. }
\CSLRightInline{Gomulkiewicz, R. \emph{et al.} Dos and don'ts of testing
the geographic mosaic theory of coevolution. \emph{Heredity}
\textbf{98}, 249--258 (2007).}

\leavevmode\hypertarget{ref-Thrall2007CoeSym}{}%
\CSLLeftMargin{12. }
\CSLRightInline{Thrall, P. H., Hochberg, M. E., Burdon, J. J. \& Bever,
J. D. Coevolution of symbiotic mutualists and parasites in a community
context. \emph{Trends in Ecology \& Evolution} \textbf{22}, 120--126
(2007).}

\leavevmode\hypertarget{ref-Drexler2014EcoEvo}{}%
\CSLLeftMargin{13. }
\CSLRightInline{Drexler, J. F., Corman, V. M. \& Drosten, C. Ecology,
evolution and classification of bat coronaviruses in the aftermath of
SARS. \emph{Antiviral Research} \textbf{101}, 45--56 (2014).}

\leavevmode\hypertarget{ref-Calisher2006BatImp}{}%
\CSLLeftMargin{14. }
\CSLRightInline{Calisher, C. H., Childs, J. E., Field, H. E., Holmes, K.
V. \& Schountz, T. Bats: Important Reservoir Hosts of Emerging Viruses.
\emph{Clinical Microbiology Reviews} \textbf{19}, 531--545 (2006).}

\leavevmode\hypertarget{ref-Moratelli2015BatZoo}{}%
\CSLLeftMargin{15. }
\CSLRightInline{Moratelli, R. \& Calisher, C. H. Bats and zoonotic
viruses: Can we confidently link bats with emerging deadly viruses?
\emph{Memórias do Instituto Oswaldo Cruz} \textbf{110}, 1--22 (2015).}

\leavevmode\hypertarget{ref-Mollentze2020VirZoo}{}%
\CSLLeftMargin{16. }
\CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
homogenous among taxonomic orders of mammalian and avian reservoir
hosts. \emph{Proceedings of the National Academy of Sciences}
\textbf{117}, 9423--9430 (2020).}

\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
\CSLLeftMargin{17. }
\CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
cross-species viral transmission risk. \emph{Nature} 1--1 (2022)
doi:\href{https://doi.org/10.1038/s41586-022-04788-w}{10.1038/s41586-022-04788-w}.}

\leavevmode\hypertarget{ref-Springer2013PhyBat}{}%
\CSLLeftMargin{18. }
\CSLRightInline{Springer, M. S. Phylogenetics: Bats United, Microbats
Divided. \emph{Current Biology} \textbf{23}, R999--R1001 (2013).}

\leavevmode\hypertarget{ref-Teeling2005MolPhy}{}%
\CSLLeftMargin{19. }
\CSLRightInline{Teeling, E. C. \emph{et al.} A Molecular Phylogeny for
Bats Illuminates Biogeography and the Fossil Record. \emph{Science (New
York, N.Y.)} \textbf{307}, 580--584 (2005).}

\leavevmode\hypertarget{ref-Villalobos2010DivFie}{}%
\CSLLeftMargin{20. }
\CSLRightInline{Villalobos, F. \& Arita, H. T. The diversity field of
New World leaf-nosed bats (Phyllostomidae). \emph{Global Ecology and
Biogeography} \textbf{19}, 200--211 (2010).}

\leavevmode\hypertarget{ref-Becker2022OptPre}{}%
\CSLLeftMargin{21. }
\CSLRightInline{Becker, D. J. \emph{et al.} Optimising predictive models
to prioritise viral discovery in zoonotic reservoirs. \emph{The Lancet
Microbe} (2022)
doi:\href{https://doi.org/10.1016/S2666-5247(21)00245-7}{10.1016/S2666-5247(21)00245-7}.}

\leavevmode\hypertarget{ref-Allen2017GloHot}{}%
\CSLLeftMargin{22. }
\CSLRightInline{Allen, T. \emph{et al.} Global hotspots and correlates
of emerging zoonotic diseases. \emph{Nature Communications} \textbf{8},
(2017).}

\leavevmode\hypertarget{ref-Olival2017HosVir}{}%
\CSLLeftMargin{23. }
\CSLRightInline{Olival, K. J. \emph{et al.} Host and viral traits
predict zoonotic spillover from mammals. \emph{Nature} \textbf{546},
646--650 (2017).}

\leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
\CSLLeftMargin{24. }
\CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
virome diversity in bats. \emph{Current Opinion in Virology}
\textbf{52}, 192--202 (2022).}

\leavevmode\hypertarget{ref-Letko2020BatVir}{}%
\CSLLeftMargin{25. }
\CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
K. \& Munster, V. J. Bat-borne virus diversity, spillover and emergence.
\emph{Nature Reviews Microbiology} \textbf{18}, 461--471 (2020).}

\leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
\CSLLeftMargin{26. }
\CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
Systems of Bats. \emph{Frontiers in Immunology} \textbf{11}, 26 (2020).}

\leavevmode\hypertarget{ref-Gorbunova2020WorGoe}{}%
\CSLLeftMargin{27. }
\CSLRightInline{Gorbunova, V., Seluanov, A. \& Kennedy, B. K. The World
Goes Bats: Living Longer and Tolerating Viruses. \emph{Cell Metabolism}
\textbf{32}, 31--43 (2020).}

\leavevmode\hypertarget{ref-Irving2021LesHos}{}%
\CSLLeftMargin{28. }
\CSLRightInline{Irving, A. T., Ahn, M., Goh, G., Anderson, D. E. \&
Wang, L.-F. Lessons from the host defences of bats, a unique viral
reservoir. \emph{Nature} \textbf{589}, 363--370 (2021).}

\leavevmode\hypertarget{ref-Jebb2020SixRef}{}%
\CSLLeftMargin{29. }
\CSLRightInline{Jebb, D. \emph{et al.} Six reference-quality genomes
reveal evolution of bat adaptations. \emph{Nature} \textbf{583},
578--584 (2020).}

\leavevmode\hypertarget{ref-MorenoSantillan2021LarGen}{}%
\CSLLeftMargin{30. }
\CSLRightInline{Moreno Santillán, D. D. \emph{et al.} Large-scale genome
sampling reveals unique immunity and metabolic adaptations in bats.
\emph{Molecular Ecology} mec.16027 (2021)
doi:\href{https://doi.org/10.1111/mec.16027}{10.1111/mec.16027}.}

\leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
\CSLLeftMargin{31. }
\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
distribution of bat hosts of sarbecoviruses: Implications for
conservation and public health. \emph{Proceedings of the Royal Society
B: Biological Sciences} \textbf{289}, 20220397 (2022).}

\leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
\CSLLeftMargin{32. }
\CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
SARS-Related Coronavirus Infection in Humans, China. \emph{Virologica
Sinica} \textbf{33}, 104--107 (2018).}

\leavevmode\hypertarget{ref-Ramshaw2019DatGeo}{}%
\CSLLeftMargin{33. }
\CSLRightInline{Ramshaw, R. E. \emph{et al.} A database of geopositioned
Middle East Respiratory Syndrome Coronavirus occurrences.
\emph{Scientific Data} \textbf{6}, 318 (2019).}

\leavevmode\hypertarget{ref-Lednicky2021IsoNov}{}%
\CSLLeftMargin{34. }
\CSLRightInline{Lednicky, J. A. \emph{et al.} Isolation of a Novel
Recombinant Canine Coronavirus from a Visitor to Haiti: Further Evidence
of Transmission of Coronaviruses of Zoonotic Origin to Humans.
\emph{Clinical Infectious Diseases: An Official Publication of the
Infectious Diseases Society of America} ciab924 (2021)
doi:\href{https://doi.org/10.1093/cid/ciab924}{10.1093/cid/ciab924}.}

\leavevmode\hypertarget{ref-Vlasova2022AniAlp}{}%
\CSLLeftMargin{35. }
\CSLRightInline{Vlasova, A. N. \emph{et al.} Animal alphacoronaviruses
found in human patients with acute respiratory illness in different
countries. \emph{Emerging Microbes \& Infections} \textbf{11}, 699--702
(2022).}

\leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
\CSLLeftMargin{36. }
\CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
Sister Species of Eidolon (Pteropodidae) with Evidence for Widespread
Panmixia. \emph{Acta Chiropterologica} \textbf{16}, 279--292 (2014).}

\leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
\CSLLeftMargin{37. }
\CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
Sequences From Malagasy Fruit Bats Define a Unique Evolutionary History
for This Coronavirus Clade. \emph{Frontiers in Public Health}
\textbf{10}, (2022).}

\leavevmode\hypertarget{ref-Hazarie2021IntPop}{}%
\CSLLeftMargin{38. }
\CSLRightInline{Hazarie, S., Soriano-Paños, D., Arenas, A.,
Gómez-Gardeñes, J. \& Ghoshal, G. Interplay between population density
and mobility in determining the spread of epidemics in cities.
\emph{Communications Physics} \textbf{4}, 191 (2021).}

\leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
\CSLLeftMargin{39. }
\CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
Origin in China. \emph{Emerging Infectious Diseases} \textbf{10},
1030--1037 (2004).}

\leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
\CSLLeftMargin{40. }
\CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
Becker, D. J. Sampling strategies and pre-pandemic surveillance gaps for
bat coronaviruses. 2022.06.15.496296 (2022)
doi:\href{https://doi.org/10.1101/2022.06.15.496296}{10.1101/2022.06.15.496296}.}

\leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
\CSLLeftMargin{41. }
\CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
synthesis of ecological and evolutionary determinants of bat diversity
across spatial scales. \emph{BMC ecology} \textbf{18}, 18 (2018).}
